---
title: "Case Study: Parsing JSON"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{json}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

_WIP: At this point this vignette is only a rough draft._

```{r setup}
library(parsy)
```

## JSON Null

Parsing a JSON null value is just parsing the literal string `null`. Since we
are not building the parser from character parsers, there's little to do but
parse the string with our fundamental parser `parse_string()` and transform the
result into a `NULL` with `parser_map()`.

```{r}
parse_json_null <- parse_string("null") |> parser_map(\(.) NULL)
parse_json_null |> parser_run("null")
```

An interesting side-effect of the way parser state in {parsy} is implemented via
attributes is that an R `NULL` value shows up as a lack of the `value` attribute
on a successful parse result.

## JSON Boolean

Parsing a boolean is slightly more complex as it requires making a choice. A
boolean in JSON is either `true` or `false`. We use the alternative parser
combinator `%|%` to check both values with their own string parsers. The resulting
parser can have as a result value either the string `"true"` or `"false"` which
we `switch()` on to return the corresponding logical value in R.

```{r}
parse_json_bool <- parse_string("true") %|% parse_string("false") |>
  parser_map(\(x) switch(x, true = TRUE, false = FALSE))

parse_json_bool |> parser_run("true")
parse_json_bool |> parser_run("false")
```

A slight tangent: The logical operator `|` is generic and it could be natural
and convenient to implement the parser combinator `%|%` as a method for it.
However, taking that route would have two major downsides:

* `|` is an _S4 generic_ that dispatches on both arguments.
* `|` has lower operator precedence than the base forward pipe operator `|>`.

The first point means we lose the benefits of lazy evaluation; all combined
parsers would have to be executed in order for method dispatch to resolve.
Our `%|%` instead dispatches only on the argument on the left hand side and
evaluates the right hand side lazily, allowing short-circuiting.

The second point means that in order to pipe the combined parser we would have
to write `(p | q) |> f()`. Custom operators like `%|%` have higher precedence
than `|>` and result in the nicer form of `p %|% q  |> f()`.

## JSON Number

Moving on to numbers, things start to get a bit more complex. At its core we
parse some digits, but we'll also need to deal with:

* Leading zeroes for parsing natural numbers; JSON does not allow them.
* The `-` symbol to extend our parser to integers.
* The `.` decimal separator to extend to floating point values.

We have the requisite tools available, but it is at this point where we have
to acknowledge the interpreted and vector-based nature of R. Rather than writing
out the combinations requiring several single-character parsers, it becomes
practical to reach out for regular expressions with `parse_regexp()`.

<!-- TODO: Just 0 -->

To parse an integer we match at the start of input (`^`) an optional minus sign
(`-?`) followed by a non-zero digit (`[1-9]`) and 0 or more digits (`[0-9]*`).
We also delegate the interpretation of the minus sign and the conversion of
a string of digits into a base 10 number to R's `as.integer()`.

```{r}
parse_integer <- parse_regexp("^-?[1-9][0-9]*") |> parser_map(as.integer)

parse_integer |> parser_run("42")
parse_integer |> parser_run("-1")
```

We can parse integers succesfully. Let's also observe the parser failing with
leading zeroes:

```{r}
parse_integer |> parser_run("042")
```

<!-- TODO: 0.1 -->

Extending the regular expression to allow for an optional decimal separator `.`
followed by some digits is straightforward. As JSON does not distinguish between
integers and doubles, we also change over to `as.double()` to always convert the
result into a floating point value in R.

``` {r}
parse_json_number <- parse_regexp("^-?[1-9][0-9]*\\.?[0-9]*") |> parser_map(as.double)
parse_json_number |> parser_run("19.5")
```

The extension to also parse scientific notation (e.g. `1e5`) is left as an
exercise to the reader.

## JSON String

In JSON a string is a bunch of characters enclosed by double quotes.
